Road Accident Prediction in Time and Space

### 2019-04-20 paper/paper.tex
Update results with correction of negative examples with 0.5 sample of 1/1000 dataset

\subsection{Positive and negative examples generation}
This machine learning problem can be stated as a binary classification problem, the positive class is the occurrence of an accident and the negative class is the non-occurrence of an accident on a given road at a given date and hour. The vehicle collision dataset contains 150 thousands of collisions, among them $134 489$ contain the date, the hour and the localization of the accident. For each of these accidents, we identify the closest road segment. These pairs of a time and a road segment are used as positive examples. For the negative examples, we generated a random sample of 5 millions combinations of time and road segment which did not appear in the collision dataset.  The generation of these examples together with the join operation with the data from the other datasets made our dataset generation heavy. It resulted in an impractical amount of time and memory space requirements to generate the dataset. Our first implementation was querying the Historical Climate Data API in real-time with a cache mechanism. Our idea was to collect only the weather stations and hours necessary for our sample of negative examples, but it resulted in bad performances. We got a performance increase by first building a Spark data frame with all the Historical Climate Data for weather stations around Montreal and then joining the two datasets. We conducted a detailed analysis of our algorithm to improve its performances. We notably obtained a good performance increase by not persisting intermediary results of the road segment identification for accidents. As opposed to what we initially thought, recomputing these results was faster than writing and reading them in the cache. Finally, the identification of the road segment corresponding to accidents was very memory intensive, we modified this step to be executed by batch of one month.  With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we managed to reduce the processing time to a reasonable time. We also used a cluster from “Compute Canada” to take maximum advantage of Apache Spark distributed nature for the generation of examples and the hyper-parameter tuning of our models.

@@

\section{Results}
The training time of the machine learning algorithm is very long due to the size of our dataset and the latest improvement in term of feature engineering have been discovered very late. Therefore, we currently only have the results with the Balanced Random Forest using the latest features. The results from the previous training of other algorithm are not comparable since they do not use the same features. We will add the missing results in later versions of this document.

These results have been obtained by training the Balanced Random Forest algorithm on the whole dataset of positive samples, but with only a fifth of the dataset of negative examples we have generated. Therefore the data imbalance is reduced to a factor of 8. According to our previous experiments, the results are almost the same when using the full dataset of negative examples, sometimes slightly better. To evaluate this model we used a test set containing the last two years of our dataset, the model is trained on the 4 previous years and use only data from these years. For example, the "count\_accident" feature contains only the count of accidents on road segments which occurred from 2012 to 2016.

We obtain an area under the precision-recall curve of 79\% and an area under the ROC curve of 97\%.
This results in the following precision and recall values depending on the chosen probability threshold:

@@

As we can see, with a threshold set to 0.63, we can achieve a recall of 85\% with a precision of 0.52\%. With a threshold of 0.06, the model can actually detect almost all the accidents, 94\% of them, with a precision of 37\%, this corresponds to the high-risk roads and time in which we predict a risk of 37\% of accidents.

\section{Discussion}
The overall results are good but mostly rely on the count of previous accidents on the road segment. This is not an issue for accident prediction, but it does not help to understand why these roads are particularly dangerous. Accident prediction is a very hard machine learning problem because even if the risk is very high for an accident to occur, the driver might manage to avoid it in time and avoid the accident although the conditions were almost exactly the same as when an accident occurs. We would need information on the state of the drivers and the state of the vehicle to significantly improve performances.

However, we believe these results can be improved even without this important feature by adding more spatial features. We can see that one of the most important variables are the localization of the accident, but the ability to learn a spatial distribution by the random forest algorithm is limited. Therefore we believe providing more context information on where the road is located will greatly help. In another experiment, we have seen the population density and the distance to the closest billboard used as features successfully. However, it will be hard to reach good performances because of the nature of the problem without the count of accidents.

>>>>

\subsection{Positive and negative examples generation}
This machine learning problem can be stated as a binary classification problem, the positive class is the occurrence of an accident and the negative class is the non-occurrence of an accident on a given road at a given date and hour. The vehicle collision dataset contains 150 thousands of collisions, among them $134 489$ contain the date, the hour and the localization of the accident. For each of these accidents, we identify the closest road segment. These pairs of a time and a road segment are used as positive examples. For the negative examples, we generated a random sample of 2 millions combinations of time and road segment which did not appear in the collision dataset.  The generation of these examples together with the join operation with the data from the other datasets made our dataset generation heavy. It resulted in an impractical amount of time and memory space requirements to generate the dataset. Our first implementation was querying the Historical Climate Data API in real-time with a cache mechanism. Our idea was to collect only the weather stations and hours necessary for our sample of negative examples, but it resulted in bad performances. We got a performance increase by first building a Spark data frame with all the Historical Climate Data for weather stations around Montreal and then joining the two datasets. We conducted a detailed analysis of our algorithm to improve its performances. We notably obtained a good performance increase by not persisting intermediary results of the road segment identification for accidents. As opposed to what we initially thought, recomputing these results was faster than writing and reading them in the cache. Finally, the identification of the road segment corresponding to accidents was very memory intensive, we modified this step to be executed by batch of one month.  With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we managed to reduce the processing time to a reasonable time. We also used a cluster from “Compute Canada” to take maximum advantage of Apache Spark distributed nature for the generation of examples and the hyper-parameter tuning of our models.

@@

\section{Results}
These results have been obtained by training the Balanced Random Forest algorithm on the whole dataset of positive samples, but with only half of the dataset of negative examples we have generated. Therefore the data imbalance is reduced to a factor of 8. To evaluate this model we used a test set containing the last two years of our dataset, the model is trained on the 4 previous years and use only data from these years. For example, the "count\_accident" feature contains only the count of accidents on road segments which occurred from 2012 to 2016.

We obtain an area under the precision-recall curve of 54\% and an area under the ROC curve of 89\%.
This results in the following precision and recall values depending on the chosen probability threshold:

@@

As we can see, with a threshold set to 0.63, we can achieve a recall of 85\% with a precision of 0.29\%. With a threshold of 0.05, the model can actually detect almost all the accidents, 94\% of them, with a precision of 19\%, this corresponds to the high-risk roads and time.

\section{Discussion}
The overall results are good but mostly rely on the count of previous accidents on the road segment. This is not an issue for accident prediction, but it does not help to understand why these roads are particularly dangerous. Accident prediction is a very hard machine learning problem because even if the risk is very high for an accident to occur, and even if an accident almost occur, the by reacting quickly, the driver can manage to avoid it although the conditions were almost exactly the same as when an accident occurs. We would need information on the state of the drivers and the state of the vehicle to significantly improve performances.

However, we believe these results can be improved even without this important feature by adding more spatial features. We can see that the most important variables are the features of the road segment. Therefore we believe providing more context information on where the road is located will greatly help. In another experiment, we have seen the population density and the distance to the closest billboard used as features successfully. However, even with more features and more data, it will be hard to reach very good performances because of the nature of the problem.
